---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---
```{r setoption, echo=FALSE}
library(knitr)
opts_chunk$set(echo = TRUE, results = "asis", cache = TRUE)
```

### Loding important libraries
```{r}
library(dplyr)
library(ggplot2)
```


## Loading and preprocessing the data
Load Processed/transformed data (if necessary) into a format suitable for your analysis.
```{r Loading and preprocessing the data}
df <- read.csv("activity.csv", na.strings = "NA", header = T)
df$date <- as.Date(as.character(df$date)) # convert date column type.
```


## What is mean total number of steps taken per day?
1. Make a histogram of the total number of steps taken each day
```{r Histogram of total steps}
daily <- df %>% group_by(date) %>% na.omit() %>%
    summarise(totalSteps = sum(steps))
qplot(daily$totalSteps, bins = 30, xlab = "number of steps",
      main = "total number of steps taken per day")
```

2. Calculate the total number of steps taken per day
```{r mean and median}
mean(daily$totalSteps)
median(daily$totalSteps)
```


## What is the average daily activity pattern?
1. Make a time series plot
```{r time series plot}
intervally <- df %>% group_by(interval) %>% na.omit() %>%
    summarise(avgSteps = mean(steps))
qplot(x = interval, y = avgSteps, data = intervally, geom = 'line',
      ylab = "average steps",
      main = "The average daily activity pattern")
```

2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r}
intervally %>% filter(avgSteps == max(avgSteps))
```



## Imputing missing values
1. Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)
```{r}
sum(is.na(df))
```

2. Create a new dataset that is equal to the original dataset but with the missing data filled in.
```{r}
imputedDF <- df %>% replace(is.na(df), mean(df$steps, na.rm = T))
sum(is.na(imputedDF))
```

3. Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day.
```{r}
daily_imputed <- imputedDF %>% group_by(date) %>% na.omit() %>%
    summarise(totalSteps = sum(steps))
qplot(daily_imputed$totalSteps, bins = 30, xlab = "number of steps",
      main = "total number of steps taken per day")
```

4. Compare the mean and median of Old and New data
```{r mean}
old_mean <- mean(daily$totalSteps)
print(old_mean)
new_mean <- mean(daily_imputed$totalSteps)
print(new_mean)
```
```{r median}
old_median <- median(daily$totalSteps)
print(old_median)
new_median <- median(daily_imputed$totalSteps)
print(new_median)
```


## Are there differences in activity patterns between weekdays and weekends?
1. Create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.
```{r}
imputedDF$days <- weekdays(imputedDF$date)
imputedDF$days <- ifelse(imputedDF$days=="Saturday" | imputedDF$days=="Sunday", "weekend", "weekday")
table(imputedDF$days)
```

2. Make a panel plot containing a time series plot of the 5-minute interval
```{r}
imputedDF <- imputedDF %>% group_by(interval, days) %>%
    summarise(avgSteps = mean(steps))
ggplot(data = imputedDF, aes(x=interval, y=avgSteps, color=days)) + 
    geom_line() + facet_grid(days ~ .) + xlab("interval") +
    ylab("average steps") +
    ggtitle("Average number of steps taken in weekdays or weekends")
```
